\documentclass[palatino]{epflnotes}

\title{Computer simulation of physical systems}
\author{Guillermo Julián Moreno}
\date{16/17 - Fall semester}

% Additional packages

% --------------------

\begin{document}
\frontmatter
\pagestyle{plain}
\maketitle

\tableofcontents
\mainmatter
% Content

\chapter{Preliminaries}

\textit{Trigger warning: Mathematics explained by a physics professor.}

\section{Differential equations}

We are interested in a general differential equation of the form \[ \od{y}{x} = f(x,y) \], that is, first order equations. We are not interested in second order problems, although we can always downgrade. For example, given a momentum equation such as \[ m \od[2]{z}{t} = F(z) \], we can simplify to two first-order equation such as \[ \od{z}{t} = \frac{p(t)}{m} \qquad \od{p}{t} = F(z) \]

All this boils down that we're are going to have initial value problems for $\vy(0) = y_0$, and then we will want to get how does $\vy$ evolve with time.

In order to solve these kinds of equations numerically, we will have several points? steps?. The first one is \textbf{discretization}, introducing evaluations at a finite number of points with step $h$\footnote{That is, points given as $x_n = x_0 + n·h$.}. The issue is the size of the step, which depends on the application being done.

The second point is a \textbf{recursion formula}: that is, how to get the $n$-th value $y_n$ as a function of the previous values $y_{n-1}, y_{n-2}, \dotsc$.

The simplest formula is called the Euler method.

\begin{prop}[Euler method] \label{prop:EulerMethod} Let $y(x)$ be a ¿? function with the differential equation $\iod{y}{x} = f(x,y)$ with $f$ being another ¿? function. Let $y(x_n) = y_n$ be a value of the equation. Then we can obtain the next value by the given formula: \( y_{n+1} = y_n + h · f(x_n, y_n) + \mathcal{O}(h^2) \)
\end{prop}

However, this is a simple method and not very accurate, as it has a low convergence rate. Another method is based on Taylor expansion of again functions in some absolute random space

\begin{prop}[Taylor expansion method] In the same conditions (that is, no conditions) of \fref{prop:EulerMethod}, we can use a method of higher-order accuracy: \( y_{n+1} = y_n + hf + \frac{1}{2}h^2 \left(\dpd{f}{x} + f \dpd{f}{y}\right) + \mathcal{O}(h^3) \label{eq:TaylorExpansion} \)
\end{prop}

\chapter{Molecular dynamics}

\chapter{Monte Carlo simulations}

\appendix

\chapter{Evaluation methods}

There are two ways to get to the exam.

\paragraph{Choose a simulation project (Recommended)} Choose a project from the web (password \textit{simphys}). Follow two hours per week of the course for a general background. We get to the exam by handing in at the end of the semester (21\textsuperscript{st} December). Ten minutes of presentation of the project at the exam plus questions. Deadline for project approval is in two weeks (aka first week of October).

The project must be done with a person of the Faculty of Basic Sciences, in C/Fortran. It should be original research. The description must be one A4 page with the proposant (professor proposing, with name and address), the motivation behind the project, the methods envisioned/considered for the project, and the student's email and name. This has to be submitted through the webpage.

Two hard copies of the report, handed in in the professor's office or in the letterbox. It also has to be uploaded to Moodle, along with the code.

The report must have the following structure:

\begin{enumerate}
\item Introduction \& motivation. Why is the simulation is necessary and ffs it's like every other introduction.
\item Description of the model.
\item Implementation of the computer code.
\item Benchmarks.
\item Results + discussions.
\item Conclusions.
\end{enumerate}

The evaluation criteria is the following:

\begin{enumerate}
\item Choice of topic and level of challenge.
\item Achieved results: quality of the computational work and physical understanding of the problem.
\item Quality of the reports.
\item Quality of the oral presentation.
\item Answers to the questions (depth of understanding).
\item Expert present at the exam gives a feeling of the work.
\end{enumerate}

\paragraph{Exercise sessions} They give code, we do magic with the code. Then we have a report and then an exam with questions on the course.

%\chapter{---}
%\input{tex/ComputerSimulationOfPhysicalSystemsI_Exerc.tex}

\backmatter
\printindex
\end{document}
