% -*- root: ../NumericalApproximationofPDEs.tex -*-
\section{Weak formulation}

\begin{problem} Consider the following boundary value problem: \[ \begin{cases}
- (αu')' + βu' + γu = f & x ∈ Ω = (a,b) \\
u(a) = u(b) = 0
\end{cases}\] with $a, β, γ ∈ ℝ^+$ and $\appl{f}{(a,b)}{ℝ}$ a sufficiently regular function. Choose a suitable test function space $V$ and write the weak form of the problem.

\solution

We multiply by a test function $v ∈ V$, and we want to get rid of the second derivative in the term $-(αu')'$. To do that, we integrate by parts:
\[ - \int_Ω α u'' v \dif x = - \eval[1]{αu'v}_{x = a}^b + \int_Ω αu'v' \dif x \] and we choose $V = H_0^1(Ω)$ to make the first term go to zero. The weak form has then the following forms:
\begin{align*}
a(u,v) &= \int_Ω αu'v' + βu'v + γuv \dif x \\
F(v) &= \int_Ω f v \dif x
\end{align*}
\end{problem}

\begin{problem} \label{ex:ODE:WeakForm2} Choose a suitable test function space and write the weak formulation of the following problems in a weak formulation. In every case, $α, β, γ ∈ ℝ^+$.

\ppart \[ \begin{cases}
- (αu')' + βu' + γu = f & x ∈ (0,1) \\
u(0) = 0 \\
u'(1) = 3
\end{cases}\]

\ppart \[ \begin{cases}
- (αu')' + βu' + γu = f & x ∈ (0,1) \\
u(0) = 0 \\
c_1 u'(1) + c_2 u(1) = 3
\end{cases}\] with $c_1, c_2$ non-null constants.

\solution

\spart

We use the space $V = \set{v ∈ H^1(0,1) \tq v(0) = 0}$. As in the previous exercise, to remove the second derivative in the first term of the integral we have
\[ - \int_0^1 α u'' v \dif x = - \eval[1]{αu'v}_{x = 0}^1 + \int_0^1 αu'v' \dif x = - 3 α v(1) + \int_0^1 αu'v' \] and so the forms are
\begin{align*}
a(u,v) &= \int_0^1 αu'v' + βu'v + γuv \dif x \\
F(v) &= \int_0^1 f v \dif x + 3 α v(1)
\end{align*}

\spart

We use again the space $V = \set{v ∈ H^1(0,1) \tq v(0) = 0}$ which comes restricted by the Dirichlet condition on 0. Then, we see what happens in the first term:
\[ - \int_0^1 α u'' v \dif x = - \eval[1]{αu'v}_{x = 0}^1 + \int_0^1 αu'v' \dif x = - α \frac{1 - c_2 u(1)}{c_1} v(1) + \int_0^1 αu'v' \] and so the forms are
\begin{align*}
a(u,v) &= - α \frac{c_2 u(1)}{c_1} v(1) + \int_0^1 αu'v' + βu'v + γuv \dif x \\
F(v) &= \int_0^1 f v \dif x + α \frac{1}{c_1} v(1)
\end{align*}

\end{problem}

\begin{problem} Consider the following boundary value problem, with non-homogeneous boundary conditions: \[ \begin{cases}
-(αu')' = f & x ∈ (0,1) \\
u(0) = 1 \\
u(1) = 2
\end{cases} \] with $α ∈ ℝ^+$.

Use the lifting technique to write a weak formulation of this problem.

\solution

By what we have seen in the previous exercises, the space of test functions is $H_0^1$ and the forms are \begin{align*}
a(u,v) &= \int_0^1 α u' v' \dif x \\
F(v) &= \int_0^1 f v \dif x
\end{align*} although the issue here is that $u ∉ H_0^1$. However, we can decompose $u = \tilde{u} + R$ where $\tilde{u} ∈ H_0^1$ and $R(0) = u(0)$, $R(1) = u(1)$ and in that case the weak formulation changes to \begin{align*}
a(u,v) &= F(v) \\
a(\tilde{u} + R, v) &= F(v) \\
a(\tilde{u}, v) &= F(v) - a(R, v) ≝ \tilde{F}(v)
\end{align*}

Here, choosing $R = 1 + x$ is enough and the modified form $\tilde{F}$ is \[ \tilde{F}(v) = F(v) - a(R, v) = \int_0^1 fv \dif x - \int_0^1 αv' \dif x \]
\end{problem}

\section{Lax-Milgram theorem}

\begin{problem} Consider the following boundary value problem \[
\begin{cases}
-(μu')' +βu' + γu = f &\text{in } Ω = (a,b) \\
u(a) = 0 \\
u'(b) = g_b
\end{cases}\] with $μ, β, γ ∈ L^∞(Ω)$, $β' ∈ L^∞(Ω)$, $g_b ∈ ℝ$ and $f ∈ L^2$.

\ppart Write the problem in weak formulation.
\ppart Give sufficient conditions on $μ, β, γ$ such that the problem has an unique solution by the \nref{thm:Theory:LaxMilgram}.

\solution

\spart

Recalling \fref{ex:ODE:WeakForm2}, the weak form with mixed Neumann-Dirichlet boundary conditions is \begin{align*}
a(u,v) &= \int_0^1 μu'v' + βu'v + γuv \dif x \\
F(v) &= \int_0^1 f v \dif x + g_b  μ(1) v(1)
\end{align*} with $V = \set{v ∈ H^1(Ω) \tq v(0) = 0}$.

\spart

Let's prove continuity of $a, F$ and coercivity of $F$. For continuity of $a$, assuming $M_μ = \essup μ$ and similarly for $β,γ$ we have
\begin{align*}
\abs{a(u,v)} ≤ \abs{\int_0^1 M_μ u'v' + M_β + M_γ uv \dif x} ≤ M_μ \abs{u}_{H^1}  \abs{v}_{H^1} + β \abs{u}_{H^1}\norm{v}_{L^2} + M_γ \norm{u}_{L^2} \norm{v}_{L^2}
\end{align*} which is bounded by the correct norms when applying \nref{thm:Fund:PoincareInequality}.

For continuity of $F$, we see that $g_b μ(1) v(1) = g_b μ(1) \left(\int_{∂Ω} v^2\right)^{\sfrac{1}{2}}$ and so we can bound it by the \nref{thm:Fund:Trace} \[ F(v) = \int_0^1 fv \dif x + g_b μ(1) v(1) ≤ \left(\norm{f}_L^2 + C_T g_b μ(1)\right) \norm{v}_{V}\]

For coercivity of $a$, we need $m_μ = \esinf μ > 0$ and we have it.

\end{problem}

\begin{problem} The temperature $T$ of a bar of length $L$ and constant section $A$ satisfies, for $x ∈ (0,L)$, the following boundary value problem:
\[
\begin{cases}
- k AT'' + σpT = 0 & x ∈ (0,L) \\
T(0) = T_0 \\
T'(L) = 0
\end{cases}\] with $p$ the perimeter of the section $A$, $k$ the termal conductivity, $σ$ the convection coefficient and $T_0$ is a given value.

\ppart Compute the analytical solution of the problem.
\ppart Write the weak formulation of the problem.

\solution

\spart

The analytical solution must fulfill \[ \frac{T''}{T} = \frac{σp}{kA} ≝ λ^2 > 0 \] so that its general solution is given by $T(x) = C_1 e^{λx} + C_2 e^{-λx}$. By replacing on the boundary conditions, we get the constants:
\begin{align*}
T'(L) &= 0 \\
λ(C_1 e^{λL} - C_2 e^{-λL}) &= 0 \\
C_1 e^{λL} &= C_2 e^{-λL} \\
C_1 &= C_2 e^{-2λL} \\
\end{align*} and
\begin{align*}
T(0) &= T_0 \\
C_1 + C_2 &= T_0 \\
C_2(e^{-2λL} + 1) &= T_0 \\
C_2 &= \frac{T_0}{e^{-2λL} + 1} \\
C_1 &= \frac{T_0 e^{-2λL}}{e^{-2λL} + 1}
\end{align*}

\spart

Done already. Use a lifting function.

\end{problem}

\section{GLS Stabilization}

\begin{problem} Consider the following advection-diffusion-reaction problem in non-conservative form.
\[
\begin{cases}
Lu ≝ -(μu')' + bu' + γu = 0 & \text{in } Ω = (0,1) \\
u(0) = 0 \\
(μu')(1) = 0
\end{cases}\] with $μ,b, γ ∈ L^∞(Ω)$ and $f ∈ L^2(Ω)$.

\ppart Write the weak formulation of the problem.
\ppart Show that the problem has a unique solution.
\ppart Write the Galerkin formulation of the problem with GLS stabilization.
\ppart Show that under the same hypothesis of the second part of the problem, the following inequality holds for the GLS stabilized problem: \[ \norm{u_h}_{GLS} ≤ C_0 \norm{f}_{L^2} \] where $C_0$ is a constant and \[ \norm{v}_{\text{GLS}}^2 = \norm{\sqrt{μ}v'}^2_{L^2} + \norm{\sqrt{\bar{γ}} v}^2_{L^2} + \sum_{K ∈ \mesh} \pesc{Lv, τ_K Lv} \] with \[ \bar{γ} = -\frac{1}{2} b'(x) + γ(x) \]

\solution

\spart

As usual, multiply by a test function $v ∈ V = \set{v ∈ H^1(Ω) \tq v(0) = 0}$ and integrate by parts to get rid of the second derivative:
\[ - \int_Ω (μu')' v \dif x = - \eval[1]{v (μu') \dif x}_{x = 0}^1+ \int_Ω μu'v' \dif x = \int_Ω μu'v' \dif x \]

Thus, the weak formulation is given by the forms
\begin{align*}
a(u,v) &= \int_Ω μu'v' + bu' + γu \dif x \\
F(v) &= \int_Ω fv \dif x
\end{align*}

\spart

I am not going to prove yet another time again the hypothesis of Lax-Milgram on these forms. Do it yourself.

\spart

The problem under GLS stabilization\footnote{See \fref{sec:ODEs:StronglyConsistentMethods} for more details.} implies finding $u_h ∈ V_h$ such that for any $v_h ∈ V_h$ we have \[ a(u_h, v_h) + \linop(u_h, v_h) = F(v_h) + ψ(v_h)\] with \begin{align*}
\linop(u_h, v_h) &= \sum_{K ∈ \mesh} \pesc{Lu_h, τ_K Lv_h} \\
Ψ(v_h) &= \sum_{K ∈ \mesh} \pesc{f, τ_K Lv_h}
\end{align*} and $τ_K$ a function given by \[ τ_K(x) = δ \frac{h_K}{\abs{b}(x)} \qquad δ > 0 \]

\end{problem}

\section{Parabolic problems}

\begin{problem} Consider the following parabolic problem in $Ω = (0,1) ⊂ ℝ$: Find $\appl{u}{Ω ×ℝ^+}{ℝ}$ such that \[\begin{cases} \dpd{u}{t} - \dpd{}{x} \left(μ \dpd{u}{x} - βu\right) + γ u = 0 & x ∈ Ω × ℝ^+ \\
u(x,0) = u_0(x) & x ∈ Ω\\
u(0,t) = η & t > 0 \\
μ(1) \dpd{u}{x} (1,t) - β(1) u(1,t) = -δu(1,t) & t > 0
\end{cases} \] where $μ, β, β', γ ∈ L^∞(0,1)$, $μ(x) ≥ μ_0 > 0$, $u_0 \equiv u_0(x) ∈ L^2(Ω)$ are given functions, and $η ∈ ℝ$ and $δ ∈ ℝ^+$ are constants.

\ppart Write the weak formulation of the problem, study coercivity and continuity of the bilinear form $a(·, ·)$.

\ppart Show that the Galerkin formulation is stable

\solution

\spart

As usual, multiply by test function $v ∈ V = \set{v ∈ H^1(Ω) \st v(0) = 0}$ and integrate by parts:
\begin{align*}
\dpd{u}{t} - \dpd{}{x} \left(μ \dpd{u}{x} - βu\right) + γ u &= \pesc{∂_t u, v} - \int_Ω \dpd{}{x} \left(μ \dpd{u}{x} - βu\right)v + γ u v \dif x = \\
&= \pesc{∂_t u, v} - \int_{∂Ω} \left(μ \dpd{u}{x} - βu\right)v \dif x + \int_Ω \left(μ \dpd{u}{x} - βu\right) v_x \dif x + \int_Ω γ u v \dif x \\
&= \pesc{∂_t u, v} + \underbracket{δu(1,t) v(1,t) + \int_Ω \left(μ \dpd{u}{x} - βu\right) v_x \dif x + \int_Ω γ u v \dif x}_{a(u,v)}
\end{align*}

We need a lifting function, so we say $u = \tilde{u} + R$ with $\tilde{u} ∈ V$, so $F(v) = - a(R, v)$. To fulfill boundary conditions, we say $R = η(1 - x)$ and then \[ F(v) = -a(R,v) = η \int_Ω \left(-μ - β(1-x)\right)v_x + γ(1-x)v \dif x \]

Continuity of $a$:
\begin{align*}
\abs{a(u,v)} &≤ δ \abs{u(1,t) v(1,t)} + \abs{\int_Ω μ u_x v_x \dif x} + \abs{\int_Ω βuv_x \dif x} + \abs{\int_Ω γuv \dif x} \\
	&≤ δC_{T,u} C_{T,v} \norm{u}_{V} \norm{v}_{V} + M_μ C_{P,u} C_{P,v} \norm{u}_V \norm{v}_V + M_β C_{P,v} \norm{u}_V \norm{v}_{V} + M_γ \norm{u}_V \norm{v}_V = \\
	&= (∂C_T + M_μC_P + M_βC_P +M_γ)\norm{u}_V \norm{v}_V
\end{align*} where $M_μ = \essup \abs{μ} = \norm{μ}_{∞} < ∞$ because  $μ ∈ L^∞$, similarly with others, and we used the \nref{thm:Fund:Trace} and \nref{thm:Fund:PoincareInequality} for inequalities on the norms of the border and seminorms.

For coercivity, using that $uu_x = \frac{1}{2} ∂_x (u^2)$ and noting $\bar{γ} = \sfrac{β'}{2} + γ$:
\begin{align*}
a(u,u) &= δ u(1,t)^2 + \int_Ω μu_x^2 \dif x - \int_{Ω} β uu_x \dif x + \int_Ω γu^2 \dif x \\
	&≥ δ u(1,t)^2 + μ_0 \abs{u}_V^2 - \frac{1}{2}\int_Ω β ∂_x (u^2)+ β'u^2 \dif x  + \frac{1}{2}\int_{Ω} β' u^2 \dif x + \int_Ω γu^2 \dif x \\
	&= δ u(1,t)^2 + μ_0 \abs{u}_V^2 - \frac{1}{2} \int_Ω ∂_x(βu^2) \dif x + \int_Ω \left(\frac{β'}{2} + γ\right)u^2 \dif x \\
	&≥ δu(1,t)^2 + μ_0 \abs{u}_V^2 - \frac{1}{2} \eval[1]{βu^2}_{x = 0}^1 + m_{\bar{γ}} \norm{u}^2_{L^2} \\
	&≥ \left(δ - \frac{β(1)}{2}\right)u(1,t)^2 + μ_0\abs{u}^2_V+ m_{\bar{γ}} \norm{u}_{L^2}^2
\end{align*} so if $δ - \sfrac{β(1)}{2} > 0$ and $m_{\bar{γ}} = \esinf \bar{γ} ≥ 0$ we have coercivity with $α = μ_0$.

\spart

Say $v_h = u_h$, so then \begin{align*}
\pesc{∂_tu_h, u_h} + a(u_h, u_h) &= F(u_h) \\
\int_Ω \dpd{u_h}{t} u_h + μ_0 \norm{u_h}^2_{V} &≥ C \norm{u_h}_{V} \\
\frac{1}{2}\int_Ω \dpd{u_h^2}{t} + μ_0 \norm{u_h}^2_V &≥ C \norm{u_h}_V \\
\frac{1}{2} \dpd{}{t} \norm{u_h}^2_{L^2} + μ_0 \norm{u_h}^2_V &≥ C \norm{u_h}_V \\
\frac{1}{2} \norm{u_h(t)}_{L^2}^2 - \frac{1}{2}\norm{u_{h0}}_{L^2}^2 + μ_0 \norm{u_h}^2_{V} &≥ C \norm{u_h}_{L^2}
\end{align*}

\end{problem}
